---
title: "Lr finder dev notebook"
output: github_document
---


# Resnext with R
Notebook to writte the freeze and unfreeze function for the fastai wrapper.

```{r}
# This R environment comes with many helpful analytics packages installed
# It is defined by the kaggle/rstats Docker image: https://github.com/kaggle/docker-rstats
# For example, here's a helpful package to load

library(tidyverse) # metapackage of all tidyverse packages

# Input data files are available in the read-only "../input/" directory
# For example, running this (by clicking run or pressing Shift+Enter) will list all files under the input directory

list.files(path = "../input")

```

```{r}
#devtools::install_github("henry090/fastai",dependencies=FALSE)
#fastai::install_fastai(gpu = TRUE)
```

```{r}
#fastai::install_fastai(gpu = TRUE)
```

```{r}
library(fastai)
```

## Data loader

I am using this two ressources : [the documentation of fastai](https://docs.fast.ai/vision.data.html#ImageDataLoaders.from_df) and the [tutorial of the wrapper](https://henry090.github.io/fastai/articles/basic_img_class.html).

```{r}
path_img = 'cassava-leaf-disease-classification/train_images/'
```

```{r}
#library(data.table)
```

```{r}
labels<-read_csv('cassava-leaf-disease-classification//train.csv')
head(labels)
```

```{r}
dataloader <- fastai::ImageDataLoaders_from_df(df=labels, path=path_img, bs=64, seed=6, device = 'cuda', num_workers=0, item_tfms = RandomResizedCrop(224, min_scale=0.75))
```

num_workers=0 is mandatory to not have the error "RuntimeError: DataLoader worker (pid(s) 482) exited unexpectedly".

```{r}
dataloader %>% show_batch()
```

If I had to bet, it is something related to kaggle/R environnement asking to define paremeters, as it works well in [the tutorial](https://henry090.github.io/fastai/articles/basic_img_class.html).

```{r}
learnR <- dataloader %>% cnn_learner(xresnet18(), metrics = accuracy,  model_dir="fastai_model/") #prettier
```

```{r}
#learnR %>% freeze()
#Error in freeze(.): could not find function "freeze"
```

### Control of the manipulation

```{r}
learnR$freeze()
learnR$summary()
```

```{r}
learnR$unfreeze()
learnR$summary()
```

### New functions

```{r}
learnR %>% freeze()

learnR$summary()
```


At the moment xresnet50() create out of memory errors.

```{r}
learnR %>% lr_find()
learnR %>% plot_lr_find(dpi = 200)
```


```{r}
learnR %>% fit_one_cycle(n_epoch = 1) 
```


```{r}
learnR %>% unfreeze()
```

```{r}
learnR$summary()
```

```{r}
learnR %>% lr_find()
learnR %>% plot_lr_find(dpi = 200)
```


